Three Problems for Sophistication

### 2015-01-19 Facticity/Facticity/paper.tex
aant lezen

\newcommand{\M}{\mathscr M}
\newcommand{\C}{\mathscr C}
\newcommand{\T}{\mathscr T}
\newcommand{\F}{\mathscr F}
\renewcommand{\P}{\mathscr P}
\newcommand{\K}{\mathscr K}
\newcommand{\X}{\mathscr X}
\newcommand{\B}{\mathbb B}
\newcommand{\D}{\Delta}
\newcommand{\N}{\mathbb N}
\newcommand{\tn}[1]{\textnormal{#1}}
\newcommand{\pair}[1]{\left\langle{#1}\right\rangle}
\newcommand{\concat}{\oplus}
\newcommand{\symb}[1]{\texttt{#1}}
\newcommand{\br}[1]{\overline{#1}}
\newcommand{\s}{S}
\newcommand{\num}[1]{#1_1, #1_2, \ldots}

\newcommand{\phiv}{\mathbf\phi}
\newcommand{\psiv}{\mathbf\psi}

@@

% --- DELETE BEFORE SUBMUSSIONS ---
\pagestyle{headings}

@@

@@

There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would call ``complex''. For example, a sequence of a million coin flips is exceedingly unlikely to have a Kolmogorov complexity of much less than one million bits, even though there is nothing complex about flipping a lot of coins---it's just hard work. For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data. This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed, see the related work section below. In this paper, we investigate three problems with these definitions, and show that they are unsatisfactory. We believe the problem is fundamental, and provide several intuitive arguments why these attempts to define sophistication must fail.

@@

The Kolmogorov complexity of a binary string $x$ is, roughly, the length of the shortest computer program that will output $x$. This length depends on which programming language is used, but only by a constant, independent of $x$. This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings. We can see it as \emph{a property of the data}.

@@

Underfitting (Section~\ref{section:underfitting}) is the opposite problem, where the selected model is simple, but fails to represent part of the structure in the data. It is an especially serious problem for sophistication because the models for used in sophistication are so powerful. In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language. Such interpreters, or \emph{universal models}, can be described in a constant amount of bits, and can generate the data with from a program the length of the Kolmogorov complexity, essentially encoding all information as noise. If complex models are penalized to avoid overfitting, then the problem becomes to make sure that universal models are not \emph{always} preferred for complex data. The usual workaround is to restrict the set of allowed models to exclude the universal ones. For example, the model class may be restricted to the set of all total recursive functions (only programs that halt for all inputs) rather than all partial recursive functions (all programs). While this patches the problem of universal models to an extent, it is questionable whether it adequately solves the problem of underfitting in general.

Finally, in the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as unlikely to capture all structure in the data, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable models of the data generating mechanism. Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that sometimes the data allows for multiple equivalent interpretations, and that there is no such thing as sophistication.

@@

Koppel \cite{koppelSoph1988,koppel1991almost} also built on Kolmogorov and Cover's work to define a measure of the meaningful information in a binary string, called \emph{sophistication}, the term we adopt in this paper. He changed the class of models from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity, to be able to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors. We will only consider the finite version (based on regular functions) in this paper.

@@

Antunes et al. \cite{antunes2009sophistication} note that the constant cutoff used by Koppel to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability, a result in the spirit of this paper. They propose a different definition called \emph{coarse sophistication}. As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth. Vereshchagin \cite{vereshchagin2013algorithmic} proposes an alternative solution to the instability issue.

@@

Let $\B = \{0,1\}^*$. We deal with partial functions $\phi: \B \to \B$, which we will also call \emph{models}. Such functions are \emph{computable} if they can be computed by a Turing machine as described in \cite[Definition~1.7.1]{li1993introduction}. A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix set. A function $f$ is \emph{total} if $D_f = \B$. A prefix-free function is \emph{complete} if every infinite binary string has a member of $D_f$ as a prefix. \footnotemark

@@

\begin{definition}[Numberings]
A numbering is an enumeration of of the partial computable functions, denoted as $\phi_1, \phi_2, \ldots$. We fix one canonical, which we know to be an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $\phi_i(y)$, where the computation diverges if $\phi_i(y) = \infty$. We call a numbering $\num{\psi}$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ such that for all $i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{b(i)}$. We use the greek letters without a subscript (eg. $\phi$ and $\psi$) to refer to the numbering as a whole.
\end{definition}

\begin{definition}[Prefix encoding function]
  A prefix encoding function $f:\B\to\B$ is an injective
  function that maps the binary strings to a prefix-free set. There
  are prefix encoding functions $g$ with the property that
  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
  denote it by $\br{x}=g(x)$.
\end{definition}

\begin{definition}[Complexity]\belowdisplayskip=-12pt
Let $\phi_1,\phi_2,\ldots$ be the canonical acceptable enumeration of $\C$. Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponds to the set of all computable prefix functions. We define the plain and prefix Kolmogorov complexity as 
\begin{align*}
C(x)&=\min\{|\bar\imath y|:\phi_i(y)=x\},\\
K(x)&=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}.\\
\end{align*}\label{definition:complexity}\vspace{-18pt}
\end{definition}
This definition of $K$ is uncommon. It is equivalent to the standard definition as used in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).

@@

\begin{definition}[Models and model classes] A \emph{model class} is a set of models. We distinguish four important model classes:
  \begin{itemize}
  \item $\C$ is the set of all partial computable functions,
  \item $\K\subset\C$ is the set of all prefix-free
    functions,
  \item $\T\subset\C$ is the set of all total functions,
  \item $\P\subset\K$ is the set of all complete, prefix-free functions,
  \item $\F\subset\P$ contains, for every finite set $S$, a computable
    surjective function mapping the binary sequences of length
    $\lceil\log S\rceil$ onto $S$. In other words, this is the model class of finite sets.
  \end{itemize}
\end{definition}
We note that classes $\T$ and $\P$ are unusual in that they are not enumerable.
\begin{definition}[Representations and code-lengths]
A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model and $m(y)=x$.
Define the following code-length functions for the representations:
\begin{itemize}
\item $L(m)$ is the length of a self-delimiting description of
  the model $m$, i.e. it is the length function of a prefix code for models. 
\item $L^m(x)=L(m)+\min\{|y|:m(y)=x\}$ measures the number of bits required to describe both the model $m$ and the data $x$ using the model.
\item $L^\M(x)=\min\{L^m(x):m\in\M\}$ measures the number of bits
  required to describe $x$ using the best model in the model class.
\end{itemize}
\end{definition}
We can now define which of these representations express the data most efficiently:
\begin{definition}[Candidate set]
The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
\[
  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}
\]
with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $\tau^\circ(x)=L^\M(x)$. We will often omit the subscript if the default threshold function is used. 
\end{definition}

\begin{definition}[Sophistication]\belowdisplayskip=-12pt
  The model complexity, meaningful information, effective complexity, facticity or \emph{sophistication},  of a string $x$ given a model class $\M$ and threshold function $\tau$ is
  \[
  \s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}\p
  \]\label{definition:sophistication}
\end{definition}
Finally, we must formalize what we mean when we say that we expect sophistication to be invariant. It may be that a less strict invariance than we have for the Kolmogorov complexity, still yields a useful sophistication.
\begin{definition}
Let $\gamma:\N\to \N$ be some nondecreasing function. Two functions $f$ and $g$ are $\gamma$-equivalent if $|f(x) - g(x)| \in O(\gamma(\max(f(x), g(x)))$.
\end{definition}
We say that a formalization of sophistication $\s_\phi$ built on acceptable numbering $\phi$ is $\gamma$-invariant to the choice of numbering is $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$. 
\begin{lemma}
If a function $\S_\phi(x)$ is bounded for a given numbering, it is either not $\gamma$-invariant or it is  $\gamma$-equal to a bounded function.
\end{lemma}
The proof follows directly from the definitions. Most lemmas in the rest of the paper will simply prove that a given variant is bounded for a specific numbering, implicitly using this lemma to show that the variant is unsound.

\section{Inefficient indices}
\label{section:indices}
We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and an input. Can we assume that, when such a description is as short as the Kolmogorov complexity, it automatically creates a good separation into the structural part of the data and the noise? As we will show, the representation of function plays a fundamental part in answering this question.

The issues can be very subtle and require a careful attention that is rarely given in the literature. We will first illustrate what happens if any acceptable numbering is allowed:

\begin{lemma}
Let $\phi$ be some acceptable numbering, and $\s_\phi(x)$ be any function which gives the length of some index $i$ such that $\phi_i(y) = x$ for some $y$.

No such function can be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
\end{lemma}
\begin{proof}
Choose some $\gamma(n) \in o(n)$. Assume towards contradiction that the sophistication is $\gamma$-invariant and unbounded.

Let $\num{\phi}$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.

We now define two new acceptable numberings $\num{\psi}$ and $\num{\xi}$:
\[
\psi_j(x) = \begin{cases}
	\phi_i(x) \;\text{if}\; j = s_{2i}\\
	\infty\;\text{otherwise}
\end{cases}\,\,\,
\xi_j(x) = 
\begin{cases}
	\phi_i(x) \;\text{if}\; j = s_{2i+1} \\
	\infty \;\text{otherwise}\p
\end{cases}
\]
Let $\s_\psi$ and $\s_\xi$ be the sophistication functions built on these enumerations. Choose any $x$ and assume w.l.o.g. that $\s_\psi(x) < \s_\xi(x)$. By construction of the numberings $\psi$ and $\xi$, we have $2\;\s_\psi(x)  \leq \s_\xi(x)$, which gives us:
\begin{align*}
	\s_\xi(x) - \s_\psi(x) &\geq \frac{1}{2}\, \s_\xi(x)  \in \Omega(\s_\xi(x))\p 
\end{align*}
Since, by assumption, $\s_\xi$ can be arbitrarily large and $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, we have a contradiction.
\end{proof}
In this proof, we deliberately used ridiculously inefficient index functions. Obviously, one could simply pick a more efficient index; the point is that many definitions of sophistication do not mention any such restriction. As we shall see, any inefficiency in the index will lead to a lack of invariance in sophistication.

The treatments that fall prey to this problem are Koppel and Atlan's \cite{koppelSoph1988,koppel1991almost},  and several of its variants (\cite{antunes2009sophistication,antunes2013sophistication}).

@@

\begin{definition}\belowdisplayskip=-12pt
  The complexity of a partial recursive function is defined by
  \begin{align*}
    C_\phi(f) &:= \min\{C(i):\phi_i=f\}\;\tn{,}\\
    K_\phi(f) &:= \min\{K(i):\phi_i=f\}\p
  \end{align*} \label{definition:model-complexity}
\end{definition}
These definitions also appear in \cite{grunwald2004shannon} and \cite{vitanyi2004meaningful}. We will show that these values are invariant up to a constant (ie. complexity is an objective property of a function):
\begin{lemma}[Invariance]
Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be any two acceptable numberings. There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function such that $\psi_i$ to $\phi_{g(i)}$.
\begin{align*}
K_\phi(f) &= \min\left\{ K_\phi(i) : \phi_i= f\right\} 
\geq \min\left\{ K_\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K_\psi(i) : \psi_{g(i)}= f\right\} - c
\geq \min\left\{ K_\psi(i) : \psi_i= f\right\} - c = K_\psi(f).
\end{align*}
The opposite inequality can be achieved by reversing $\phi$ and $\psi$. 
\end{proof}
This lemma shows that the complexity of a function does not depend on the choice of
enumeration by more than a constant term.

There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature,
\begin{enumerate}
  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used. Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$. We construct such an index below. This approach is taken by Adriaans \cite{adriaans2012facticity}.
  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data. With this approach, the total code length is $K(\phi_i)+|y|$, where $y$ is an input to the model $\phi_i$ such that $\phi_i(y)=x$. The difference is that in this approach, we can construct everything on the basis of a single acceptable numbering, without additional constraints. This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
\end{enumerate}
To show that the two perspective are equivalent, we must show that there exists an acceptable numbering which is efficient enough.

\begin{definition}[Faithful Numbering]\label{def:faithful}
  A numbering $\psi_1,\psi_2,\ldots$ of the partial recursive
  functions is \emph{faithful} if there is a constant $c$ such that
  for all indices $i$ there is a $j$ such that $\psi_i=\psi_j$ and
  $|j|\le C(\psi_j)+c$.
\end{definition}
This definition is a refinement of the one used in \cite{adriaans2012facticity}. \footnotemark

@@

To show that $\psi$ is acceptable, let $\phi_j$ denote the identity
function. Then a $\phi$-index $i$ can be mapped to a $\psi$-index
using the computable function $r(i)=\bar\jmath i$, so that
$\psi_{r(i)}(y)=\psi_{\bar\jmath i}(y)=\phi_i(y)$. For the reverse,
define $\phi_v(\bar\imath p, y)=\phi_{\phi_i(p)}(y)$. For fixed
$\bar\imath p$, the 
$s^n_m$-theorem \cite{kleene193notation} states that we can compute the $h$
such that $\phi_h(y)=\phi_v(\bar\imath p,y)$. Let $h(\bar\imath p)$
denote this index as a function of the program; further define
$h(q)=\tn{div}$ if $q$ cannot be expressed as $\bar\imath p$. By
construction $h$ is total and computable. To check that the mapping
returns the correct function, rewrite $\phi_{h(\bar\imath
  p)}(y)=\phi_v(\bar\imath p,y)=\phi_{\phi_i(p)}(y)=\psi_{\bar\imath p}(y)$.
\end{proof}
Unfortunately, even with a faithful index, we can fail if the prefix-function used to separate the model from its input is inefficient:
\begin{lemma}
Let $L(\phi_i) = |\bar\imath|$ for a faithful index, and some prefix encoding function, and let $\s^\C$ be defined on this $L$ with the default threshold function.

>>>>

\newcommand{\M}{\mathscr M}
\newcommand{\C}{\mathscr C}
\newcommand{\T}{\mathscr T}
\newcommand{\F}{\mathscr F}
\renewcommand{\P}{\mathscr P}
\newcommand{\K}{\mathscr K}
\newcommand{\X}{\mathscr X}
\newcommand{\B}{\mathbb B}
\newcommand{\D}{\Delta}
\newcommand{\N}{\mathbb N}
\newcommand{\tn}[1]{\textnormal{#1}}
\newcommand{\pair}[1]{\left\langle{#1}\right\rangle}
\newcommand{\concat}{\oplus}
\newcommand{\symb}[1]{\texttt{#1}}
\newcommand{\br}[1]{\overline{#1}}
\newcommand{\s}{S}
\newcommand{\dom}[1]{\mathop{\tn{dom}(#1)}}

@@

% --- DELETE BEFORE SUBMISSIONS ---
\pagestyle{headings}

@@

\sdr{Course sophistication handled sufficiently?}
\sdr{Do we really need the $\gamma$-invariance? Or can $\gamma$ be 1?}
\sdr{We need the definitions! But maybe we can simplify, do we really need the candidate set?}

@@

There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would call ``complex''. For example, a sequence of a million coin flips is exceedingly unlikely to have a Kolmogorov complexity of much less than one million bits, even though there is nothing complex about flipping a lot of coins---it's just hard work. For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data. This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed, see the related work section below. In this paper, we investigate three problems with these definitions, and show that they are unsatisfactory. We believe the problem is fundamental, and provide arguments why such attempts to define sophistication must fail.

@@

The Kolmogorov complexity of a binary string $x$, denoted $K(x)$, is (roughly) the length of the shortest computer program that will output $x$. This length depends on which programming language is used, but, as shown by a core result known as the Invariance Theorem, the language affects the result only by a constant, independent of $x$. This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings. We can see it as \emph{a property of the data}.

@@

Underfitting (Section~\ref{section:underfitting}) is the opposite problem, where the selected model is simple, but fails to represent part of the structure in the data. It is an especially serious problem for sophistication because the models for used in sophistication are so powerful. In particular, in any programming language, there are programs that implement an interpreter for \emph{another} language. Such interpreters, or \emph{universal models}, can be described in a constant amount of bits, and can generate the data with an input of the  the length of the Kolmogorov complexity, essentially encoding all information as noise. If complex models are penalized to avoid overfitting, then the problem becomes to make sure that simple universal models are not \emph{always} preferred for complex data. The usual workaround is to restrict the set of allowed models to exclude the universal ones. For example, the model class may be restricted to the set of all total recursive functions (only programs that halt for all inputs) rather than all partial recursive functions (all programs). While this patches the problem with universal models, it is questionable whether it adequately solves the problem of underfitting in general.

Finally, in the discussion in Section~\ref{section:conclusion} we argue that while two-part coding can yield useful insights into the structure of the data and allows us to identify some models as unlikely to capture all structure in the data, it is probably not possible to uniquely separate structure from noise and identify a \emph{single} model as ``best'': in general many models, of vastly different complexities, may be reasonable models of the data generating mechanism. Rather than doggedly trying to ``fix'' this property of algorithmic statistics, we propose embracing the idea that sometimes the data allows for multiple interpretations, and that there is no such thing as sophistication.

@@

Koppel \cite{koppelSoph1988,koppel1991almost} also built on Kolmogorov and Cover's work to define a measure of the meaningful information in a binary string, called \emph{sophistication}, the term we adopt in this paper. Koppel changed the class of models from finite sets to total functions and used monotone functions, instead of prefix-free Kolmogorov complexity, to be able to extend the theory to infinite strings, an approach which has been abandoned by subsequent authors. We will only consider the finite version (based on regular functions) in this paper.

@@

Antunes et al. \cite{antunes2009sophistication} note that the constant cutoff used by Koppel to separate candidate two-part descriptions from those deemed to be too inefficient, is a source of instability, a result in the spirit of this paper. They propose a different definition called \emph{coarse sophistication}. As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth. Vereshchagin \cite{vereshchagin2013algorithmic} proposes an alternative solution to the instability issue. \sdr{But if his stuff works, this paper is nonsense, so we should say more about it!}

@@

Let $\B = \{0,1\}^*$. We deal with partial functions $\phi: \B \to \B$, which we will also call \emph{models}. Such functions are \emph{computable} if they can be computed by a Turing machine as described in \cite[Definition~1.7.1]{li1993introduction}. A function is prefix-free if its domain, $\dom{f} = \{y : f(y) \neq \infty\}$ is a prefix free set. A function $f$ is \emph{total} if $\dom{f} = \B$. A prefix-free function is \emph{complete} if every infinite binary string has a member of $\dom{f}$ as a prefix.\footnotemark

\sdr{I think this footnote should not be a footnote and better explained. It's important}

@@

\begin{definition}[Numberings]
A numbering $\psi_1,\psi_2,\ldots$ is an enumeration of of the partial computable functions. We fix one canonical numbering $\phi$ that we know to be effective, that is for any $i \in \N, y\in \B$, we can effectively compute $\phi_i(y)$, where the computation diverges if $\phi_i(y) = \infty$. We call a numbering $\psi$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ such that for all $i$, $\phi_i = \psi_{b(i)}$ and  $\psi_i = \phi_{a(i)}$.  We abbreviate the sequence by its symbol without the index: $\phi=\phi_1,\phi_2,\ldots$
\end{definition}

\begin{definition}[Prefix encoding function]
  A prefix encoding function $f:\B\to\B$ is an injective
  function that maps the binary strings to a prefix-free set. There
  are prefix encoding functions $g$ with the property that
  $|g(x)|\le|x|+2\log(|x|+2)|$ \cite{li1993introduction}. We fix such a function and
  denote it using a bar, $\br{x}=g(x)$.
\end{definition}

\sdr{Clearly explain what it means if a function function is total, complete, prefix. Do we need to fuss about equating the numbers and binary strings? Perhaps it's better to do the conversions explicitly in this paper}

\begin{definition}[Complexity]\belowdisplayskip=-12pt
Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ is the set of all computable prefix functions. \sdr{Cite textbook that this is possible.} We define the plain and prefix Kolmogorov complexity as 
\begin{align*}
C(x)&=\min\{|\bar\imath y|:\phi_i(y)=x\},\\
K(x)&=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}.\\
\end{align*}\label{definition:complexity}\vspace{-18pt}
\end{definition}
This definition of $K$ is uncommon. It is equivalent to the standard definition in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).

@@

\begin{definition}[Models and model classes] A \emph{model class} is a set of models. We distinguish five important model classes:
  \begin{itemize}
  \item $\C$ is the set of all partial computable functions,
  \item $\K\subset\C$ is the set of all prefix-free
    functions,
  \item $\T\subset\C$ is the set of all total functions,
  \item $\P\subset\K$ is the set of all complete, prefix-free functions,
  \item $\F\subset\P$ contains, for every finite set $S$, a computable
    surjective function mapping the binary sequences of length
    $\lceil\log S\rceil$ onto $S$. In other words, this is the model class of finite sets.
  \end{itemize}
\end{definition}
We note that classes $\T$ and $\P$ are unusual in that they are not enumerable.
\begin{definition}[Representations and code-lengths]
A pair $(m,y)$ is a \emph{representation} for $x$ if $m$ is a model and $m(y)=x$.
Define the following code-length functions for the representations:
\begin{itemize}
\item $L(m)$ is the length of a self-delimiting description of
  the model $m$, i.e. it is the length function of a prefix code for models. 
\item $L^m(x)=L(m)+\min\{|y|:m(y)=x\}$ measures the number of bits required to describe both the model $m$ and the data $x$ using the model.
\item $L^\M(x)=\min\{L^m(x):m\in\M\}$ measures the number of bits
  required to describe $x$ using the best model in the model class.
\end{itemize}
\end{definition}
We can now define which of these representations express the data most efficiently:
\begin{definition}[Candidate set]
The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
\[
  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}
\]
with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $\tau^\circ(x)=L^\M(x)$. We will often omit the subscript if the default threshold function is used. 
\end{definition}

\begin{definition}[Sophistication]\belowdisplayskip=-12pt
  The model complexity, meaningful information, effective complexity, facticity or \emph{sophistication},  of a string $x$ given a model class $\M$ and threshold function $\tau$ is
  \[
  \s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}\p
  \]\label{definition:sophistication}
\end{definition}
Finally, we must formalize what we mean when we say that we expect sophistication to be invariant. It may be that a less strict invariance than we have for the Kolmogorov complexity, still yields a useful sophistication.
\begin{definition}
Let $\gamma:\N\to \N$ be some nondecreasing function. Two functions $f$ and $g$ are $\gamma$-equivalent if $|f(x) - g(x)| \in O(\gamma(\max(f(x), g(x)))$.
\end{definition}
We say that a formalization of sophistication $\s_\phi$ built on acceptable numbering $\phi$ is $\gamma$-invariant to the choice of numbering is $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$. 
\begin{lemma}\label{lem:broken}
Given a slack function $\gamma$, if $\s_\phi(x)$ is either $\gamma$-equivalent to $0$ or to $K(x)$ for \emph{some} numbering $\phi$, then either this equivalence exists for all other numberings as well, or $\s_\phi(x)$ is not $\gamma$-invariant.
\end{lemma}
The proof follows directly from the definitions. Most lemmas in the rest of the paper will simply prove that a given variant is bounded for a specific numbering, implicitly using this lemma to show that the variant is unsound.

\section{Inefficient indices}
\label{section:indices}
We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; the index of a computable function and an input to that function. Can we assume that, when such a description is as short as the Kolmogorov complexity, it automatically creates a good separation into the structural part of the data and the noise? As we will show, this depends in part on the used enumeration $\phi$.

The issue is subtle and require a careful attention that is rarely given in the literature. We will first illustrate what happens if any acceptable numbering is allowed:

\begin{lemma}
Any sophistication $\s_\phi(x)$ that gives the length of some index $i$ such that $\phi_i(y) = x$ for some $y$ cannot be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
\end{lemma}
\begin{proof}
Choose some $\gamma(n) \in o(n)$. Assume towards contradiction that the sophistication is $\gamma$-invariant and unbounded.

Let $\phi$ be the canonical acceptable numbering. Let $s_i$ be the binary string consisting of $2^{i}$ zeroes followed by a one.

We now define two new acceptable numberings $\psi$ and $\chi$:
\[
\psi_j(x) = \begin{cases}
	\phi_i(x) \;\text{if}\; j = s_{2i}\\
	\infty\;\text{otherwise}
\end{cases}\,\,\,
\xi_j(x) = 
\begin{cases}
	\phi_i(x) \;\text{if}\; j = s_{2i+1} \\
	\infty \;\text{otherwise}\p
\end{cases}
\]
Now choose any $x$ and assume w.l.o.g. that $\s_\psi(x) < \s_\xi(x)$. By construction of the numberings $\psi$ and $\xi$, we have $2\;\s_\psi(x)  \leq \s_\xi(x)$, which gives us:
\begin{align*}
	\s_\xi(x) - \s_\psi(x) &\geq \frac{1}{2}\, \s_\xi(x)  \in \Omega(\s_\xi(x))\p 
\end{align*}
Since, by assumption, $\s_\xi$ can be arbitrarily large and $\s_\xi(x) - \s_\psi(x) \in O(\gamma(\s_\xi(x))) \subset o(\s_\xi(x))$, we have a contradiction.
\end{proof}
In this proof, we deliberately used ridiculously inefficient index functions. Obviously, one could simply pick a more efficient index; the point is that many definitions of sophistication do not mention any such restriction. As we shall see, any inefficiency in the index will lead to a lack of invariance in sophistication.

This is a serious problem with several papers, in particular
Koppel and Atlan's \cite{koppelSoph1988,koppel1991almost} and several of its variants \cite{antunes2009sophistication,antunes2013sophistication}.

@@

\begin{definition}\belowdisplayskip=-12pt
  The complexity of a partial recursive function is defined by
  \begin{align*}
    C_\phi(f) &:= \min\{C(i):\phi_i=f\}\;\tn{,}\\
    K_\phi(f) &:= \min\{K(i):\phi_i=f\}\p
  \end{align*} \label{definition:model-complexity}
\end{definition}
These definitions also appear in \cite{grunwald2004shannon} and \cite{vitanyi2004meaningful}. We will show that these values are invariant up to a constant (ie. complexity is an objective property of a function):
\begin{lemma}[Invariance]
Let $\phi$ and $\psi$ be any two acceptable numberings. There exists a constant $c$ such that $\left| K_\phi(f) - K_\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function such that $\psi_i=\phi_{g(i)}$.
\begin{align*}
K_\phi(f) &= \min\left\{ K_\phi(i) : \phi_i= f\right\} 
\geq \min\left\{ K_\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K_\psi(i) : \psi_{g(i)}= f\right\} - c
\geq \min\left\{ K_\psi(i) : \psi_i= f\right\} - c' = K_\psi(f).
\end{align*}
The opposite inequality can be achieved by reversing $\phi$ and $\psi$. 
\end{proof}
This lemma shows that the complexity of a function does not depend on the choice of
enumeration by more than a constant term.

There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature,
\begin{enumerate}
  \item We maintain that the representation witnessing the Kolmogorov complexity determines the sophistication, but impose the additional constraint that an efficient, acceptable numbering is used. Specifically, that for every function $f$ there is an $i$ with $\phi_i=f$ and $|i|=C(f)$. We construct such an index below. This approach is taken by Adriaans \cite{adriaans2012facticity}.
  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity to measure the complexity of the \emph{model} (as in Definition~\ref{definition:model-complexity}), which is then the first part of a two-part code describing the data. With this approach, the total code length is $K(f)+|y|$, where $y$ is an input to the model $f$ such that $f(y)=x$. The difference is that in this approach, we do not need any constraints on the default numbering $\phi$. This approach is used in  \cite{cover1985kolmogorov,gacs2001algorithmic,vitanyi2004meaningful,gellmann1996information}.
\end{enumerate}
To show that the two perspectives are equivalent, we must show that there exists an acceptable numbering which is efficient enough.

\begin{definition}[Faithful Numbering]\label{def:faithful}
  A numbering $\psi$ of the partial recursive
  functions is \emph{faithful} if there is a constant $c$ such that
  for all indices $i$ there is a $j$ such that $\psi_i=\psi_j$ and
  $|j|\le C(\psi_j)+c$.
\end{definition}
This definition is a refinement of the one used in \cite{adriaans2012facticity}.\footnotemark

@@

To show that $\psi$ is acceptable, let $\phi_j$ denote the identity
function. Then a $\phi$-index $i$ can be mapped to a $\psi$-index
using the computable function $r(i)=\bar\jmath i$, so that
$\psi_{r(i)}(y)=\psi_{\bar\jmath i}(y)=\phi_i(y)$. For the reverse,
define $\phi_v(\bar\imath p, y)=\phi_{\phi_i(p)}(y)$. For fixed
$\bar\imath p$, the 
$s^n_m$-theorem \cite{kleene193notation} states that we can compute the $h$
such that $\phi_h(y)=\phi_v(\bar\imath p,y)$. Let $h(\bar\imath p)$
denote this index as a function of the program; further define
$h(q)=\tn{div}$ if $q$ cannot be expressed as $\bar\imath p$. By
construction $h$ is total and computable. To check that the mapping
returns the correct function, rewrite $\phi_{h(\bar\imath
  p)}(y)=\phi_v(\bar\imath p,y)=\phi_{\phi_i(p)}(y)=\psi_{\bar\imath p}(y)$.
\end{proof}
\sdr{Explain following better: point out that it provides a slight imbalance in favour of putting stuff in the data part, which, if the model class is $\C$, already to underfitting. Refer to the underfitting section.}
Even with a faithful index, we can fail if the prefix-function used to separate the model from its input is inefficient:
\begin{lemma}
Let $L(\phi_i) = |\bar\imath|$ for a faithful index, and some prefix encoding function, and let $\s^\C$ be defined on this $L$ with the default threshold function. \sdr{Dit wringt als een gek om t zo te zeggen, moet eleganter. Volgens mij gebruiken we maar twee $L$ functies: de index met een bar erop, en $K$.}
